### 2022-04-25 LaTeX_Plos/ResponseToReviewers.tex

\blue{Citations of these recent works have been added to the text.}

\textbf{Reviewer: 1}

\textbf{Recommendation:} This paper investigates the use of silver nanoparticle and inks incorporated into knitted medical textiles that show durable anti-viral activity, even after extended washing in bleach. 
The paper is interesting, generally well-written and organised. 
However, several issues need to be addressed before it can be considered for publication. Thus, I would recommend major changes to the manuscript before resubmission and reconsideration.

\blue{We thank the reviewer for their assessment that our paper is ``interesting, generally wel-written and organised.  The manuscript has been updated based on their comments below.}

@@

FIGURES:
- XRD spectra should have the correct x-axis labelleing. ‘Angle’ is insufficient.
- For figure 4, the SEM images are useful, but I think an inset with much higher magnification is required for the reader to ascertain if any changes to the nanoparticle loading etc., has occurred. It is not possible to do so at the 10 micron level.
- Perhaps a minor point, but the authors seem to have multiple different axis labelling styles; sometimes using brackets and other times using a ‘/’ between the label and units. A single labelling style is recommended for all figures.

\blue{The above mentioned corrections have been made. The authors choose to keep the axis labelling style. Captions have been updated to ensure clarity. }

REFERENCES:
- 33 needs correcting
%\begin{enumerate}
%\end{enumerate}

\blue{Reference 33 is formatted in the same manner as all the other references.}

@@

\newpage
\textbf{Reviewer: 2}

\textbf{Recommendation:} A reactive silver ink coating was developed to provide a bleach wash durable, repellent antiviral PET fabrics. Silver nanoparticle or reactive silver ink showed antiviral property that is very important for medical textiles. However, there are significant flaw in the present article.

\textbf{Comments:} 
1. Uniform distribution of silver on the PET fabrics is crucial to the good antiviral property. How did the authors adjust silver distribution via drop casting method? How about the silver content on the coating? Cytotoxicity is needed to characterize the coated PET fabrics.

\blue{PET samples were dampened with alcohol to improve the spreading of the silver coatings during drop casting. The silver content of the coating has been added in the Materials and Methods section. Cytotoxicity assessment is outside the scope of the work.  Line 253 in the Conclusion has been updated to state, ``Future work should analyze the antiviral mechanism, as well as the cytotoxicity of reactive silver ink treated fabrics before
commercial use."}

2. The authors need to characterize PDMS coatings, for example, the thickness of PDMS coating? How did reactive silver ink and silver nanoparticle influence the final morphology of PDMS coatings?

\blue{A description of the final morphology of samples after PDMS coatings have been added in page 4, paragraph 4 line 150 ,``The PDMS
coating appears roughened due to the underlying morphology of the silver nanoparticles
or silver ink coated fabrics." }
\commentPaul{Anything about thickness?}

3. Silver is the antiviral agent to defend against virus, so why PDMS treatment on the reactive silver ink coating had better antiviral property than silver ink coating without PDMS treatment for unwashed PET fabrics?

@@

\blue{The PDMS coating adds a repellency property to the fabric, so there is simultaneous virus deactivation and repellency. ``The PDMS coated silver 
fabrics may release less silver than without PDMS; however, a greater reduction of 
infectious virus is observed with PDMS from a combination of virus inactivation and 
liquid repellency" has been added to page 6, bottom of paragraph 2, line 224-226.}

\newpage
\textbf{Reviewer: 3}

\textbf{Recommendation:} This is a well-written article covering a nice piece of timely research, at a moment in history when there is considerable interest in antiviral textiles. The quality of the science is high and the conclusions from the article are generally very robust. However, I have a small number of concerns that should be addressed through minor revision of the MS, mainly in highlighting the limitations of the study. i.e. One might be forgiven from interpreting the work and resulting technology as being at a late stage of technical readiness so the following limitations should be transparently disclosed in the article:

\blue{We thank the Reviewer for their very positive assessment of the importance of our work as well as the quality of science and conclusions of our work.  }

@@

\textbf{Comments:} 
1. The efficacy of the reactive silver approach is clear, but are there any significant unknown risks such as toxicity or has the oxidation to elemental silver been proven to be 100\% reliable?

\blue{The ink leaves a pure silver film, so toxicity should not be different than elemental silver. The curing temperature is low and at this temperature the silver is not oxidised during curing. The XRD characterization in Figure S2 shows the silver is not oxidized on the fabric after curing. }

2. The washfastness involved single exposure experiments in one treatment regime (bleach in a sonicator) when there is little evidence provided that this is industrially relevant. Such a process provides little mechanical action, abrasion or shear forces, and certainly doesn't replicate autoclave treatments used in some areas for medical textiles. Claims are made that the technology is suitable for multiple washes and uses, but no evidence is provided to support this. Moreover, the control fabric is also lab produced - a robust demonstration of improved washfastness might have included a commercially purchased antiviral textile? In summary, far more studies would be needed before any conclusion was reached that this technology was approaching commercial viability.

\blue{Ultrasonication can produce harsh mechanical action and scrubbing via bubbles and can be industrially relevant.$^{34}$ Analysis of cytotoxicity is outside the scope of this work.  Line 253 in the Conculsion has been updated to state, 
``Future work should analyze the antiviral mechanism, as well as the cytotoxicity of reactive silver ink treated fabrics before
commercial use."}

@@

3. Breadth - medical textiles are produced in more fabric types than PET knit yet no data is presented showing that the technology is applicable on other fabric types. We should therefore acknowledge the narrowness of the study.

\blue{We appreciate the suggestion and are considering other fabrics in future work. Higher surface energy
typically results in better adhesion. PET has lower surface energy so the demonstrated procedure
should be generally applicable to most textile fibers.}
%\blue{Future work may consider other textiles and this limitation is acknowledged.  }

@@

4. Other properties - technical textiles don't just need to exhibit their desired primary benefit (in this case antiviral); does the treatment compromise any other aspects of textile properties that could be relevant to its use such as flexibility and comfort in wear?

\blue{We appreciate the suggestion and are considering analysis of flexibility and comfort 
in future work.}
%\blue{Future work may consider other textile properties and this limitation is acknowledged.  }

@@

Finally, the descriptions of materials in the Materials and Methods are too vague and this prevents the work from being easily duplicated. Trademarks such as Antocon(R) are not acknowledged which is a product brand name not a supplier. That range of wipes contains more than one PET knit wipe so the specific one used should be described. The specific bleach is not given, only that it was purchased from Giant Eagle - was it Clorox(R) or their own brand? Do they really only sell one? Could product codes be provided for other purchases from Sigma-Aldrich unless they really only sell one PBS, FBS and PDMS product?

\blue{The supplier has been added, as well as a better description of the materials and methods. The specific bleach is Giant Eagle brand and (Giant Eagle brand) has been added to the text. They only sell one. 
\commentPaul{Perhaps better description of this bleach.  Concentration level.  Sodium hypochlorite concentration level}
A better description of the other materials that were used have been changed in the Materials and Methods section. Sigma only sells one PBS and FBs. PDMS (Sylgard 184) has been added to the text. 
\commentPaul{Try to provide better description of materials even if there is only one.  Sometimes, companies will change how many they sell or change the composition of the one product they sell.}
}

@@

\newpage
\textbf{Reviewer: 4}

\textbf{Recommendation:} 
The manuscript entitled ``Reactive silver inks for antiviral, repellent medical textiles with ultrasonic bleach washing durability compared to silver nanoparticles" adequately studies and discusses the antiviral activity of silver coated PET fabrics with different silver sources in terms of wash durability. However, the manuscript seems to be unclear, making the technical and presentation aspects appear weak. The manuscript can make a good case for PLOS ONE readers with some suggested improvements. Below are issues of significant importance that should be addressed in the proposed manuscript.

@@

\textbf{Comments:}

*Fig 1 highlights the characterization .... with silver nanoparticles and inks. The authors should use terminology such ``reactive silver inks" or "silver inks” instead of inks through their manuscript.

@@

\blue{The terminology has been changed throughout the text. }

*An equal amount of silver by weight (0.2 mg) was added to each fabric. $-->$ It is better to give concentrations (mg/mL) per in2. Also, what is the solvent that authors used in drop casting method, they didn't mention the solvent even in the Materials and Methods section.

\blue{Once both coatings cure, there is no liquid remaining on the fabric; therefore, mg/mL per in$^2$ is not a valid unit of measurement of the coating on the fabric. The solvent is a proprietary blend for reactive silver ink while the solvent is sodium citrate for silver nanoparticles which is mentioned in the text. }

*Additionally, we investigate the anti-viral activity.... in solution (Fig 2c and S1 Fig) $-->$ Fig 2c is believed to be Fig 2d. Moreover, S1 Fig gives information about Adenovirus type 5 which is not mentioned anywhere in the manuscript. Did authors give antiviral activity of silver sources against HAdV5 on purpose? How did\ they comment about the silver np acting better on enveloped virus then non enveloped one?

@@

\blue{The adenovirus results have been removed for clarity. }

*Virus assay in saline are conducted by mixing silver ink or silver nanoparticles (0.1\%) in PBS...$-->$ Is 0.1\% (v/v), (w/w) or (w/v)? it is better if authors specify the type of concentration.

@@

\blue{This work has been added to the citations. }

\newpage
\textbf{Reviewer: 5}

\textbf{Recommendation:}
In the submitted manuscript, the authors showed that the silver amine complex ink, which contains reactive ionic silver, can coat PET fibres to create a low surface energy surface, superhydrophobic surface. The applied coating surface exhibits antiviral properties and is resistant against extended ultrasonic bleach washing. In the introduction section, the importance of this study has been discussed very well. The manuscript can be improved and become ready to publish if the authors can address these issues:

\blue{We thank the reviewer for acknowledging the importance of our study.}

@@

\textbf{Comments:}

53 In this study the silver amine complex is being studied, however, this ink needs to be fully chemically characterised and formulated.

\blue{The silver amine complex has been fully chemically characterised and formulated in the cited paper 39.  Additional formulation is held under intellectual property.}

@@

101 The x-axis of the XRD pattern should change to (2$\theta$ (°)). Please at Y-axis please remove numbers if it is going to be reported by Arbitrary unit or remove (a.u.).

\blue{The axes have been changed. }

104 Author mentioned that the untreated PET is fully wetted by water and Fig 2 (a) shows that the water static, contact angle, for PET is zero. Although other studies such as the Indian Journal of Fibre \& Textile Research Vol. 37, September 2012, pp. 287-291 showed the water contact angle of untreated PET fabric is 124.24$\degree$.

\blue{The paper mentioned uses weaved PET fabric which may be more repellent. The PET we used is knit and is fabricated to have high sorbency. }
\commentPaul{See above comment about PET having lower surface energy than other textiles. Perhaps, this PET is not so low surface energy?}

144. Could the authors please explain based on what parameters/reasons curing temperature sets at 150 $\degree$C for 2h for a thin layer of PDMS?
